### 2019-05-05 Rapport/TG_ITA/Bibliographic_study/conference_041818.tex
Paper PIR 2 first pages

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{verbatim} 
\usetikzlibrary{positioning}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Machine Learning Autoencoder Applied to a Binary Symmetric Channel}

\author{\IEEEauthorblockN{Eduardo Dadalto Camara Gomes}
\IEEEauthorblockA{\textit{Department of Electronics, Optronics and Signal Processing (DEOS) } \\
\textit{ISAE-SUPAERO}\\
Toulouse, France \\
Email: edadaltocg@gmail.com}}

\maketitle

\begin{abstract}

write in the end.

\end{abstract}

\begin{IEEEkeywords}
communication system, machine learning, autoencoder, binary symmetric channel, maximum a posteriori (MAP) decoder 
\end{IEEEkeywords}

\section{Introduction}
\subsection{Motivation}
Communication systems are widely available in the everyday life and are built over different structures. These structures are less or more complex; nevertheless, they share a common structure. The simpler real communication architecture defines the system as an exchange of information between two terminals through a noisy channel. As a consequence, we know the ultimate reliable data rate that can be transmitted over a noisy channel, theorized by Shannon \cite{b1} in 1948.

Since then, the research community in digital communication developed a series of algorithms to minimize bit error rate (BER) over a channel. However, the challenge of finding a \textit{efficient} solution , i.e., at same time low latency and with low error probability, for low signal to noise ratio (SNR) channels remains. Based on this vision, Tim O'Shea and Jakob Hoydis \cite{b2} pertinently remarked that traditional algorithms in the field have foundations in probability theory (e.g. maximum a posteriori (MAP) estimation used as a channel decoder), hence they are built on top of mathematically convenient models. Often, these models does not account for all the system's real imperfections, what leads to intrinsic errors, even though they are theoretically optimal.

Accordingly, machine learning (ML) algorithms does not require rigidly designed models and can take effortlessly account for non-linearities. Moreover, the design of communication systems as independently working blocks becomes obsolete, as a deep neural network (DNN) is able to actuate in an end-to-end manner. As a result, an optimized autoencoder with a stochastic layer that models channel's imperfections can substitute the block based representation. Therefore, ML based communication systems could be a better representation of realistic systems and could optimize information transmission of different blocklengths and with low decoding latency, resulting ultimately in gain of bandwidth.

Low latency and high bandwidth wireless communication are key to critical systems, such as airplanes, satellites, cellular communication and 5G operations. The latter was studied by F. D. Calabrese et al. in \cite{b3} which  demonstrated that individual based radio resource management (RRM) algorithms were outperformed by a general learning framework, resulting in significant expense reductions, while increasing performance. Thereby, ML algorithms are cardinal for future state of the art communication applications.

\subsection{Background and related work}

State of the art work in machine learning in communication systems
\subsection{Problem statement}

This research intend to implement a ML autoencoder for binary symmetric channels that outperforms the MAP decoder for a range of SNR. Using state-of-the art DNN algorithms to find the best solution of the problem, contributing to set up a higher standard in terms of performance in bit-error correction for digital communication applications. In a near future, this disruptive methodology for error correction using ML could replace mathematically optimal decoders which are the current guideline.

%\subsection{Notation}
%Special notations used throughout the paper

\section{Theoretical Background}

\subsection{Maximum a posteriori decoder}

The concept of a MAP decoder or Bahl-Cocke-Jelinek-Raviv (BCJR) algorithm is choosing the message value which maximizes the MAP probability of the corrupted information received \cite{b4}. This algorithm is known for its optimality in error correction for white noise interference. MAP decoding algorithms and it's variants Log-MAP and Max-Log-MAP are widely used as decoders for turbo codes \cite{b7}.

The MAP decoder is practically nonviable to implement \cite{b6}. Thus, for the purpose of implementation, its variants cited in the paragraph above are better suited. The Log-MAP is an optimal decoder, having equivalent performance to the MAP decoder. The Max-Log-MAP is a suboptimal decoder and will not be treated in this paper. Refer to \cite{b6} for specific details in the derivation of the Log-MAP algorithm.

Turbo decoding using Log-MAP decoder for an additive white Gaussian noise (AWGN) channel and for a binary symmetric channel (BSC) was studied in \cite{b5}. They analyzed the BER for small SNR and concluded that the results are sensible to SNR estimation of the real channel, since the estimation of this parameter is required for the metric calculation of the algorithm \cite{b6}. It means that if the real SNR is smaller, the decoder will not perform well. Hence, for channel characteristics that change over time, the application of a MAP algorithm is debatable. This conclusion serve as motivation for ML autoencoders, which could adapt to different SNR conditions, outperforming the MAP decoder for realistic applications.

\subsection{Neural network basics}

Neural networks (NN) are able to approximate any kind of function through an architecture composed of several single processing units (neurons) connected together forming a network. They are defined by \cite{b9} as are parallel distributed, learning- and self-organizing information processing systems.

In the context of learning algorithms implemented by a NN, the key issue is to find a suitable architecture that delivers the best results. To define and build this infrastructure, the activation function, the loss function and the structure must be decided.  M. Nielson \cite{b8} in his book explores a vast amount of NN architectures and how they work.

Multi-layer feed forward neural network (MLNN) as shown in Fig. \ref{fig:NN} will be treated in this work. They are known for the universal approximation property \cite{b9} and for the versatility of increasing the number of layers, creating a DNN able to undertake complex classification problems with low classification error. In more general terms, a MLNN with $L$ layers and parameter $\theta$ is a mapping of an input $\textbf{r}_0 \in \mathbb{R}^{N_0}$ to an output $\textbf{r}_L \in \mathbb{R}^{N_L}$ through $L$ iterative steps. For a fully-connected network, each layer vector is calculated in a forward propagation iterative manner, where each neuron is composed of a linear combinator and an activation function  defined as

\begin{equation}\label{eq:eqFP}
	f_{l}\left( \textbf{r}_{l-1};\theta _{l}\right) = \sigma \left( \textbf{W}_{l}\textbf{r}_{l-1}+\textbf{b}_{l}\right)
\end{equation}
where $\textbf{W}_{l}\in \mathbb{R} ^{N_{l}\times N_{l-1}}$ is the weight matrix between layers $l-1$ and $l$,  $\textbf{b}_{l}\in \mathbb{R} ^{N_l}$ is the bias vector and $\sigma$ is the activation function \cite{b2}.

In order to correct the errors from a BSC, a \textit{perceptron} processing unit is chosen to solve a classification problem for each input message. A perceptron maps a binary vector input into a single binary output digit \cite{b8}. Mathematically, it means that the function defined in \eqref{eq:eqFP}, outputs $1$ or $0$, depending on a threshold value.

For a classification problem in communication, the \textit{categorical cross-entropy} loss function $l(\textbf{p},\textbf{q}):\mathbb{R} ^{N_L}\times \mathbb{R} ^{N_L}\mapsto \mathbb{R}$ defined in \eqref{eq:c-e} is most common. It is derived from the log-likelihood of a training set where $q_j$ is the estimated probability of the outcome $j$ and $p_{j}$ is the true probability. Basically, the function measures the dissimilarity between $p_{j}$ (what you expected) and $q_{j}$ (what you obtained) \cite{b10} where $j=1,...,N_L$.
\begin{equation}\label{eq:c-e}
	l(\textbf{p},\textbf{q})=-\sum _{j}p_{j}\log \left( q_{j}\right)
\end{equation}

With all these elements set, training the neural network to calculate an accurate weight matrix $\textbf{W}$ for forward propagation requires a labeled training data set. This set is composed of pairs $ (\textbf{r}_{0,i}, \textbf{p}_i) $, where $i=1,...,S$ and $S$ is the number of training sets. It matches the input and its respective desired output. The objective is to minimize the overall loss function defined in \eqref{eq:o-c-e} in terms of the parameter $\theta$ \cite{b2}. 
\begin{equation}\label{eq:o-c-e}
L\left( \theta \right) =\dfrac {1}{S}\sum ^{S}_{i=1}l\left( \textbf{p}_{L,i},\textbf{r}_{L,i}\right)	
\end{equation}

This problem is an optimization problem and can be solved through different algorithms. For NN, an efficient way of computing this minimization is implementing the back-propagation (BP) algorithm. 	BP is classified as a supervised learning algorithm that is able to optimize a function based on some parameter and is highly parallelizable in computational terms \cite{b8}\cite{b9}.

Libraries \cite{b11} \cite{b12}.

\tikzset{%
  every neuron/.style={
    circle,
    draw,
    minimum size=0.666cm
  },
  neuron missing/.style={
    draw=none, 
    scale=2,
    text height=0.333cm,
    execute at begin node=\color{black}$\vdots$
  },
}

\begin{figure}
\centering

\begin{tikzpicture}[x=1.1cm, y=1.1cm, >=stealth]

\foreach \m/\l [count=\y] in {1,2,3,missing,4}
  \node [every neuron/.try, neuron \m/.try] (input-\m) at (0,2.5-\y) {};

\foreach \m [count=\y] in {1,missing,2}
  \node [every neuron/.try, neuron \m/.try ] (hidden-\m) at (2,2-\y*1.25) {};

\foreach \m [count=\y] in {1,missing,2}
  \node [every neuron/.try, neuron \m/.try ] (output-\m) at (4,1.5-\y) {};

\foreach \l [count=\i] in {1,2,3,m}
  \draw [<-] (input-\i) -- ++(-1,0)
    node [above, midway] {$x_\l$};

\foreach \l [count=\i] in {1,j}
  \node [above] at (hidden-\i.north) {$r_\l$};

\foreach \l [count=\i] in {1,n}
  \draw [->] (output-\i) -- ++(1,0)
    node [above, midway] {$y_\l$};

\foreach \i in {1,...,4}
  \foreach \j in {1,...,2}
    \draw [->] (input-\i) -- (hidden-\j);

\foreach \i in {1,...,2}
  \foreach \j in {1,...,2}
    \draw [->] (hidden-\i) -- (output-\j);

\foreach \l [count=\x from 0] in {Input, Hidden, Ouput}
  \node [align=center, above] at (\x*2,2) {\l \\ layer};

\end{tikzpicture}

\caption{MLNN representative diagram. Where } \label{fig:NN}
\end{figure}

\subsection{Autoencoders}

%\section{Procedure and methods}
%Explain what I Should do to achieve my goal
\begin{comment}

\subsection{Figures and Tables}

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
\centerline{\includegraphics{fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure}
\end{comment}

\begin{thebibliography}{00}
\bibitem{b1} C. E. Shannon, "A mathematical theory of communication," Bell System Technical Journal, vol. 27, no. 3, pp. 379–423, Jul. 1948.

\bibitem{b2} O'Shea, Tim \& Hoydis, Jakob, "An Introduction to Machine Learning Communications Systems", 2017.
\bibitem{b3} F. D. Calabrese, L. Wang, E. Ghadimi, G. Peters, and P. Soldati,
"Learning radio resource management in 5G networks: Framework,
opportunities and challenges," arXiv preprint arXiv:1611.10253, 2016
\bibitem{b4} W. Alexander, H. Peter \& W. Norbert, "Turbo-Decoding Without SNR Estimation",IEEE Communications Letter, vol. 4, no. 6, pp. 193-195, 2000.
\bibitem{b5} M. Jordan and R. Nichols, “The effects of channel characteristics on turbo code performance,” in Proc. Milcom’96, McLean, VA, Oct. 1996,
pp. 17–21.
\bibitem{b6} Robertson, P. , Hoeher, P. and Villebrun, E. (1997), Optimal and sub‐optimal maximum a posteriori algorithms suitable for turbo decoding. Eur. Trans. Telecomm., 8: 119-125. doi:10.1002/ett.4460080202.
\bibitem{b7} J. Hagenauer,P. Robertson, L. Papke: iterative (“Turbo”) decoding of systematic convolutional codes with the MAP and SOVA
algorithms. In: ITG-Fachbericht 130. October 1994, p. 21-29.
\bibitem{b8} M. Nielsen, "Neural Networks and Deep Learning".
\bibitem{b9} I. Mohamed, "Applications of neural networks to digital communications - a survey", 1997.
\bibitem{b10} M. Kevin, "Machine Learning: A Probabilistic Perspective", MIT, ISBN 978-0262018029, 2012.
\bibitem{b11} M. Abadi, A. Agarwal, P. Barham, et al., 
"TensorFlow: Large-scale machine learning on heterogeneous systems", 2015. Software available from tensorflow.org.
\bibitem{b12} F. Chollet, “keras,” https://github.com/fchollet/keras, 2015.

\end{thebibliography}

\end{document}