### 2019-06-25 Rapport/Presentation/Presentation/Presentation.tex
Seance 06/25/2019

\documentclass{beamer}

% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{outlines}

\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{multirow}
\usepackage{url}
%\usepackage[sorting=none]{biblatex}
\usepackage[utf8]{inputenc}
\usepackage{ifthen}
\usepackage{filecontents}
\usetikzlibrary{shapes,arrows,shadings,patterns}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\pgfplotsset{plot coordinates/math parser=false}

\makeatletter
\def\BState{\State\hskip-\ALG@thistlm}
\makeatother

\usetikzlibrary{positioning}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\makeatletter
\def\endthebibliography{%
  \def\@noitemerr{\@latex@warning{Empty `thebibliography' environment}}%
  \endlist
}
\makeatother

\makeatletter
\def\thickhline{%
  \noalign{\ifnum0=`}\fi\hrule \@height \thickarrayrulewidth \futurelet
   \reserved@a\@xthickhline}
\def\@xthickhline{\ifx\reserved@a\thickhline
               \vskip\doublerulesep
               \vskip-\thickarrayrulewidth
             \fi
      \ifnum0=`{\fi}}
\makeatother

\newlength{\thickarrayrulewidth}
\setlength{\thickarrayrulewidth}{3\arrayrulewidth}

\title[Machine Learning and Communication]{Machine Learning Autoencoder Applied to Communication Channels}

\author{E.~Dadalto Camara Gomes\inst{1}}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[] % (optional, but mostly needed)
{
  \inst{1}%
   ISAE-SUPAERO\\
   Université de Toulouse\\
    31055, Toulouse, France\\
Email: eduardo.dadalto-camara-gomes@student.isae-supaero.fr
%  \and
 % \inst{2}%
 % Department of Electronics, Optronics, and Signal processing\\
 % ISAE-SUPAERO\\
 % 31055, Toulouse, France\\
 % Email: meryem.benammar@isae-supaero.fr
 }
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

\date{ISAE-SUPAERO, 2019}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

\subject{Machine Learning and Communication}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

%\pgfdeclareimage[height=0.5cm]{university-logo}{images/ISAE}
%\logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
%\AtBeginSubsection[]
%{
%  \begin{frame}<beamer>{Outline}
%    \tableofcontents[currentsection,currentsubsection]
%  \end{frame}
%}

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Introduction}
\subsection{Context}
\begin{frame}{Context I}{Digital communication system definition}

\begin{figure}
  \centering
   \includegraphics[width=0.6\textwidth]{images/Communication_system_schema_from_book}
  \caption{Block diagram representation of a communication system.}%%\label{fig:cs}
\end{figure}

\begin{figure}
  \centering
   \includegraphics[width=0.65\textwidth]{images/simple_sys}
  \caption{Simplified communication system. Where $\textbf{u}^k$ is a source message, $\textbf{x}^n$ a code word, $\textbf{y}^n$ the received code word and $\hat{\textbf{u}}^k$ the decoded message.}\label{fig:cs}
\end{figure}

\end{frame}

\begin{frame}{Context II}{Machine learning applied to communication system}

  \begin{figure}
  \centering
   \includegraphics[width=0.35\textwidth]{images/BSC}
  \caption{Binary symmetric channel (BSC) representation.}%%\label{fig:cs}
\end{figure}

  \begin{figure}
  \centering
   \includegraphics[width=0.85\textwidth]{images/autoencoder}
  \caption{Deep neural network (DNN) based autoencoder.}%%\label{fig:cs}
\end{figure}

\end{frame}

\begin{frame}{Motivation \& Challenges}

\begin{outline}
 \1 Low latency and high bandwidth wireless communication are key to critical systems.
   \2 i.g. airplanes, satellites, cellular communication and 5G operations.
\pause
 \1 As opposite to structured algorithms, machine learning (ML) algorithms do not require rigidly designed models and can take non-linearities effortlessly into account.
\pause
 \1  ML based communication systems could be a better representation of realistic systems and could optimize information transmission of different blocklengths.
   \2 It translates to lower decoding latency, resulting ultimately in gain of bandwidth over standard methods.

\end{outline}

\end{frame}

\subsection{Problem Statement}
\begin{frame}{Problem Statement}

\begin{outline}
\1 Could ML based decoders and autoencoders for a BSC perform similarly to the MAP decoder in real applications and have lower communication delay?
\pause
\1 This work aims to contribute to set up a higher standard in terms of performance in bit-error correction and demonstrate that ML based models can reduce delay for digital communication applications. 
\end{outline}

\end{frame}

% You can reveal the parts of a slide one at a time
% with the \pause command:
%\begin{frame}{Second Slide Title}
%%  \begin{itemize}
 % \item {
%    First item.
%%    \pause % The slide will pause after showing the first item
 % }
 %% \item {   
  %  Second item.
%  }
  % You can also specify when the content should appear
  % by using <n->:
%  \item<3-> {
%    Third item.
%  }
%  \item<4-> {
%    Fourth item.
%  }
  % or you can use the \uncover command to reveal general
  % content (not just \items):
%  \item<5-> {
 %   Fifth item. \uncover<6->{Extra text in the fifth item.}
%  }
%  \end{itemize}
%\end{frame}

\section{Methodology}

\subsection{Reference Model}
\begin{frame}{Maximum a Posterior (MAP) Rule}{Implementation of a MAP decoder for a linear block code through a BSC.}

\begin{itemize}
\item{The concept of a MAP decoder algorithm for sequences is choosing a message which maximizes the a posterior probability.}
\end{itemize}

\begin{center}
$f(y) = \underset{x\in \mathcal{X}}{\text{arg max}} P_{X|Y}(x|y)$
\end{center}

 \begin{figure}[!ht]
  \centering
    \includegraphics[width=0.8\textwidth]{images/algorithm}
\end{figure}

\end{frame}

\subsection{Design \& Architecture}
\begin{frame}{Neural Network's Design and Architecture Basics}{}
%Show the architecture used for each case and remarks some important parameters
	\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.4\textwidth]{images/MLNN}
    \caption{NN representative diagram, where $\textbf{y}^n$ is the input vector, $\textbf{r}_{l}^{j}$ is a hidden layer vector
and $\hat{\textbf{u}}^{k}$ is the output vector.} \label{fig:NN}
\end{figure}
\end{frame}
	\framebreak
\begin{frame}{Array Decoder DNN Architecture and Implementation}{}
    \begin{table}
\caption{DNN array decoder architecture and training parameters. The input is a code word of size $n$ and the output is the decoded message of size $k$.} 
\begin{center}

  \begin{tabular}{ l  |  l }
    \thickhline
    \multirow{3}{*}{Decoder} & Dense: 128, activation: ReLU, input size: $n$ \\
 & Dense: 64, activation: ReLU  \\
 & Dense: 32, activation: ReLU  \\
 & Dense: $k$, activation: Sigmoid  \\ \hline
    \multicolumn{2}{l}{\textbf{Total parameters: $12776$}}\\
    \thickhline
  \end{tabular}
    \label{tab:arraydecoder}
\end{center}
\end{table}

\begin{center}
\begin{tabular}{ l  |  l | l | l  }
    \thickhline
 Loss function & Optimizer & N. Epochs & Batch Size \\ \hline
 Binary cross-entropy & Adam & $2^{16}$ & $256$ \\
    \thickhline
  \end{tabular}
    \label{tab:arraydecoderTrain}
\end{center}
\end{frame}
  \framebreak
\begin{frame}{One-hot Decoder DNN Architecture and Implementation}
    \begin{table}
\caption{DNN one-hot decoder architecture and parameters. The input is a code word of size $n$ and the output is a unique one-hot vector of size $2^k$ which is mapped into its correspondent message.} 
\begin{center}

  \begin{tabular}{ l  |  l }
    \thickhline
    \multirow{1}{*}{Decoder} & Dense: 256, activation: Softmax, input size: $n$ \\ \hline
    \multicolumn{2}{l}{\textbf{Total parameters: }$4352$}\\
    \thickhline
  \end{tabular}
    \label{tab:onehotdecoder}
\end{center}
\end{table}

\begin{center}
\begin{tabular}{ l  |  l | l | l  }
    \thickhline
 Loss function & Optimizer & N. Epochs & Batch Size \\ \hline
 Binary cross-entropy & Adam & $2^{14}$ & $256$ \\
    \thickhline
  \end{tabular}
    \label{tab:onehotdecoderTrain}
\end{center}
\end{frame}

\begin{frame}{DNN Autoencoder Design}{}
	\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.6\textwidth]{images/DNN_autoencoder}
    \caption{Representation of a DNN autoencoder composed of dense layers. Each block represents a DNN which together learn the channel encoding and decoding in an end-to-end manner. They also operate together in bit error correction.}\label{fig:DDNNAutoencoder}
\end{figure}
\end{frame}

\begin{frame}{Array Autoencoder DNN Architecture and Implementation}{This design yielded best error correction results out of 80 models tried}
    \begin{table}
\caption{DNN array autoencoder architecture and training parameters.} 
\begin{center}
  \begin{tabular}{ l  |  l }
    \thickhline
    \multirow{3}{*}{Encoder} & Dense: 512, activation: ReLU, BN\footnotemark, input size: 8 \\
 & Dense: 256, activation: ReLU, BN \\
 & Dense: 16, activation: Sigmoid  \\ \hline
    \multirow{2}{*}{Channel} & Lambda: $Round(\textbf{x})$, input size: $16$ \\
 & Lambda: $\textbf{x}\oplus \text{noise}$\\ \hline
 \multirow{2}{*}{Decoder} & Dense: 128, BN, input size: 16 \\
 & Dense: 64, activation: ReLU, BN  \\
 & Dense: 8, activation: Sigmoid  \\ \hline
    \multicolumn{2}{l}{\textbf{Total parameters: }154072}\\
    \thickhline
  \end{tabular}
    \label{tab:arrayautoencoder}
\end{center}
\end{table}

\begin{center}
\begin{tabular}{ l  |  l | l | l  }
    \thickhline
 Loss function & Optimizer & N. Epochs & Batch Size \\ \hline
 MSE & Adam & $2^{17}$ & $256$ \\
    \thickhline
  \end{tabular}
    \label{tab:arrayautoencoderTrain}
\end{center}

\footnotetext{Batch Normalization (BN)}
\end{frame}

\begin{frame}{One-hot Autoencoder DNN Architecture}{This design yielded best error correction results out of 71 models tried}
    \begin{table}
\caption{DNN one-hot autoencoder architecture and training parameters.} 
\begin{center}

  \begin{tabular}{ l  |  l }
    \thickhline
    \multirow{3}{*}{Encoder} & Dense: 196, activation: ReLU, BN, input size: 256\\
 & Dense: 128, activation: ReLU, BN  \\
 & Dense: 96, activation: ReLU, BN \\
 & Dense: 64, activation: ReLU, BN  \\
 & Dense: 32, activation: ReLU, BN  \\
 & Dense: 16, activation: Sigmoid  \\ \hline
   Channel & Lambda: $Round(\textbf{x})$, input size: $16$ \\ \hline
 Decoder & Dense: 128, activation: ReLU, BN, input size: 16 \\
 & Dense: 256, activation: Softmax  \\ \hline
    \multicolumn{2}{l}{\textbf{Total parameters: }134052}\\
    \thickhline
  \end{tabular}
    \label{tab:onehotautoencoder}
\end{center}
\end{table}

\begin{center}
\begin{tabular}{ l  |  l | l | l  }
    \thickhline
 Loss function & Optimizer & N. Epochs & Batch Size \\ \hline
 MSE & Adam & $2^{16}$ & $256$ \\
    \thickhline
  \end{tabular}
    \label{tab:onehotautoencoderTrain}
\end{center}

\end{frame}

%\subsection{Error Correction \& Predictions}
%\begin{frame}{Error Correction and Monte Carlo Simulations}{Explain how we could use NN to predict the results with certain confidence.}

%\end{frame}

%\begin{frame}{Blocks}
%\begin{block}{Block Title}
%You can also highlight sections of your presentation in a block, %%with it's own title
%\end{block}
%\begin{theorem}
%%There are separate environments for theorems, examples, definitions and proofs.
%\end{theorem}
%%\begin{example}
%Here is an example of an example block.
%\end{example}
%\end{frame}

% Placing a * after \section means it will not show in the
% outline or table of contents.

\section{Results \& Discussions}
\subsection{Deep Neural Network Based Decoders}
\begin{frame}{DNN Array Decoder Error Correction Performance}{For each of 20 different channels, 100000 messages were sent. The average error was calculated and traced. The same procedure was applied throughout the work.}

 \begin{figure}[!ht]
  \centering
    \includegraphics[width=0.7\textwidth]{images/MLNN_Mep_65536_ptrain_007}
    \caption{Array decoding BER performance. DNN trained with a channel crossover probability error of $p_t=0.07$. The decoder successfully learned the reference MAP algorithm.}\label{fig:ArrayD}
\end{figure}

\end{frame}

\begin{frame}{DNN One-hot Decoder Error Correction Performance}

   \begin{figure}[!ht]
  \centering
    \includegraphics[width=0.7\textwidth]{images/MLNN1H_Mep_16384_ptrain_0}
    \caption{One hot decoding BER performance. NN decoder trained with a channel crossover probability error of $p_t=0$. The decoder successfully learned the reference MAP algorithm.}\label{fig:1HD}
\end{figure}

\end{frame}

\subsection{Deep Neural Network Based Autoencoders}
\begin{frame}[allowframebreaks]{DNN Array Autoencoder Error Correction Performance}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.7\textwidth]{images/MAP_AutoencoderArray_Mep_65536_64_128_256_p_analysis}
    \caption{Training crossover probability simulation for the array autoencoder. $P_t=0.03$ demonstrated to have best performance to this particular architecture.}\label{fig:parrayanalysis}
\end{figure}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.7\textwidth]{images/AutoencoderArray_Mep_65536_ptrain_003_logcosh}
    \caption{Array autoencoder BER performance. DNN array autoencoder trained with a channel crossover probability error of $p_t=0.03$. The array autoencoder could not outperform the reference MAP algorithm for channels with $p<0.02$.}\label{fig:arrayautoencoder}
\end{figure}

\end{frame}

\begin{frame}{DNN One-hot Autoencoder Error Correction Performance}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.7\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
    \caption{One-hot autoencdoer BER performance. Trained without a noisy channel. No architecture could outperform the reference MAP algorithm in terms of BER.}\label{fig:1hautoencoder}
\end{figure}

\end{frame}

\subsection{Time Analysis}
\begin{frame}{Delay Time Analysis}{Calculations done by CPU}
  \begin{table}
\caption{Decoding time comparison between the reference MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the average MAP algorithm decoding time. The decoders had around a 25\% improvement in delay time in comparison to the MAP.} 
\begin{center}

  \begin{tabular}{ c | c | c}
    \thickhline
    MAP & Array Decoder & One-hot Decoder\\
    $1.00 \pm 0.02$ & $0.74 \pm 0.03$ & $0.76 \pm 0.02$ \\ \hline
    \multicolumn{2}{c|}{Array Autoencoder}  & One-hot Autoencoder\\

     \multicolumn{2}{c|}{$1.33 \pm 0.05 $} & $3.02 \pm 0.06$ \\ 
    \thickhline
  \end{tabular}
    \label{tab:timeanalysis}
\end{center}
\end{table}

\end{frame}

\section{Conclusions}

\begin{frame}{Conclusions}
  \begin{itemize}
  \item {
    The feasibility of a machine learning based channel decoder were demonstrated for a communication system with a binary symmetric channel.
  }
  \item {
    The designed decoders improved around 25\% the channel delay when compared to the maximum a posteriori rule for a $(16,8)$ code.
  }
  \item{
Their small number of parameters aligned with a GPU equipped decoding computer is promissory to be a substitute of the MAP rule and achieve same bit error correction rate in a real application. \\
  }
  \item{
We also proved that autoencoders can learn a non linear encoding function and its decoding function for a BSC.  
  }
  \item{
   There are still some challenges in training to solve.
  }
  \end{itemize}
\end{frame}

\section{Future Work}
\begin{frame}{Future Work}
  \begin{itemize}
  \item {
    A more rigorous hyper-parametric analysis could find a combination of parameters which yields better BER performance for the autoencoders.
  }
    \item{
  Using more advanced NN topology such as a recurrent neural network (RNN) or a generative adversarial network (GAN).
  }
  \item {
     Real experimental implementation of the DNN decoders could be tested to confirm their capabilities.
  }
  \end{itemize}
\end{frame}

\section*{Acknowledgment}
\begin{frame}{Acknowledgment}
  I would like to thank ...
  \begin{itemize}
   \item{The guidance and advice from my supervisor Mme. Meryem Benammar, Ph.D;}
  \item{ My colleague Rémy Zawislak who worked in the same theme and collaborated indirectly to the research;}
   \item{Mme. Marjorie Grzeskowiak Lucas, Ph.D, who through the Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO) enabled the confection of the project; and }
   \item{The evaluators, the LACS member and the audience.}
  \end{itemize}
\end{frame}

\appendix
\section<presentation>*{Bibliography}

\begin{frame}[allowframebreaks]
  \frametitle<presentation>{Bibliography}

\color{white}
\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18, DBLP:journals/corr/KeskarMNST16, DBLP:journals/corr/IoffeS15}

\bibliographystyle{ieeetr}
\bibliography{bib/ShannonCE1948,bib/CalabreseWGPS16,bib/OSheaH17,bib/Oshea2,bib/Gruber,bib/Worm,bib/Viterbi,bib/RobertsonHV97,bib/HagenauerJ,bib/JordanMA,bib/Ibnkahla,bib/Nielsen,bib/Murphy,bib/AbadiABBCCCDDDG16,bib/Keras,bib/mit_neco18_1527,bib/BenammarP18,bib/KeskarMNST16,bib/IoffeS15}

\end{frame}

\end{document}